Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-10-08 paper.tex
intro's intro fixed

    \begin{abstract} Text of abstract.  
    \end{abstract} 
    \tableofcontents
    \chapter{Introduction}
        In recent years, the volume of neuroimaging data acquired has exceeded
        both the storage and computation capacity of a standard research 
        lab workstation. With the advancement of data sharing technologies, this 
        data has been made widely available, \weird{with the only factor impeding
        research on this data being access to infrastructure and software that 
        enables efficient processing on such infrastructure}. Many research labs 
        do have access to high performance computing (HPC) clusters, however, 
        without efficient software to process this data, processing can take an 
        excessive amount of time. Ensuring efficient processing of data is 
        complex and likely beyond the knowledge of an average neuroimaging 
        researcher \tristan{Be careful with 'knowledge', some neuroimaging researchers are very compute-savvy}. As a result, frameworks for the efficient processing of 
        neuroimaging data need to be developed. \tristan{Add a sentence to transition from 'frameworks for efficient processing' to 'workflow engines'}

        Many neuroimaging workflow engines currently exist. Examples include
        Nipype~\cite{nipype}, Pipeline System for Octave and Matlab (PSOM)~\cite{10.3389/fninf.2012.00007}, LONI~\cite{REX20031033}, SPM~\cite{spm}, FastR~\cite{10.3389/fict.2016.00015},
        Automatic Analysis (AA)~\cite{10.3389/fninf.2014.00090} and Pydpiper~\cite{10.3389/fninf.2014.00067}. These 
        workflow engines aim to satisfy four criteria: 1) straightforward \tristan{easy? facilitated?} 
        workflow composition, 2) performance, 3) portability of the workflows 
        and 4) reproducibility of the analyses. While these workflow engines
        do tackle performance, it is mainly limited to minimizing computation
        time and does not consider data transfer times -- which are typically
        costly in Big Data settings. Therefore, these engines need to be adapted
        for the processing of neuroimaging data.

        Big Data frameworks differ from those available in neuroimaging as their
        focus is mainly on workflow composition and performance. Big Data 
        frameworks such as Apache Spark and Dask \tristan{Add references} both aim to be easy-to-use,
        applicable to a wide variety of analyses and improve performance in the 
        processing of Big Data. These types of systems were primarily designed 
        to be efficient on commodity infrastructure in which transferring data
        over a network would be extremely costly.

        Currently, there exists no Big Data framework which is adapted for the
        processing of neuroimaging data. However, tools such as the Thunder 
        Project \note{add citation} allow for the processing of time-series
        neuroimaging data using Spark as the backend framework. \weird{Tools such as 
        these} provide built-in functions to process the data. For users to 
        extend functionality, it is necessary for them to understand the backend
        framework. Other scientific workflow engines, such as Toil and Pegasus
        \note{add citations}, have incorporated features that enable more 
        performant workflow executions for scientific Big Data. They also enable
        the integration of Big Data frameworks as sub-pipelines within their 
        workflows. Despite integration with Big Data frameworks, it is not
        possible to create Big Data workflows using their API exclusively.

@@

        \tristan{Add a sentence to present the goal of the report.}
        In section~\ref{datasets}, we will briefly introduce the two 
        types of datasets that are giving rise to the neuroimaging big 
        data. Section~\ref{platforms} will provide a brief introduction 
        to the data-sharing and processing platforms available to 
        researchers. We will conclude our 
        introduction on the differences between commodity and HPC infrastructure
        in section~\ref{infrastructure}.

        As we wish to evaluate and compare scientific workflow frameworks with
        Big Data frameworks for the processing of neuroimaging big data, we will
        break down our analyses by the four principal features necessary for
        effective neuroimaging workflow engines. The workflow engines discussed
        will include SPM, LONI, Nipype, PSOM, FastR, AA, Pydpiper 
        for neuroimaging; Taverna, Pegasus, Make, OpenMOLE, Nextflow, Galaxy, 
        Toil for other scientific workflows; and MapReduce, Spark and Dask for
        Big Data Frameworks. Chapter \ref{workcomp} will evaluate these 
        workflows on the basis of workflow composition: we will % 'in other words' is on my list of expressions to avoid
        look at the programming languages necessary to construct these workflows
        , how easy is it to add new tools and will these workflows be understood
        by the community. In Chapter \ref{performance}, we will proceed to 
        investigate the strategies employed by these frameworks to improve the 
        performance of their workflows. Chapter \ref{portability} will discuss
        the portability of the workflows written using these frameworks. 
        Workflow reproducibility will be then discussed in Chapter 
        \ref{reproducibility}.

@@

            \subsection{input and output definitions}
                TBD

>>>>

    \begin{abstract} Text of abstract.  
    \end{abstract} 
    \tableofcontents
    \chapter{Introduction}
        In recent years, the volume of neuroimaging data acquired has exceeded
        both the storage and computation capacity of a standard research 
        lab workstation. With the advancement of data sharing technologies, 
        this data has been made widely available, however, accessibility to 
        such data remains limited due to infrastructural demands and lack of 
        performant software. Even with powerful infrastructure, such as High
        Performance Computing (HPC) infrastructure, software configured to 
        efficiently use resources is required to make execution time manageable
        . Ensuring efficient processing of data is 
        complex and limits accessibility to researchers with little programming 
        knowledge. Existing neuroimaging workflow engines primarily rely on HPC
        resource managers for performance and are therefore not well suited for
        BigData processing. As a result, frameworks for the efficient 
        processing of Big neuroimaging Data need to be developed\tristan{Add a sentence to transition from 'frameworks for efficient processing' to 'workflow engines'}

        Many neuroimaging workflow engines currently exist. Examples include
        Nipype~\cite{nipype}, Pipeline System for Octave and Matlab 
        (PSOM)~\cite{10.3389/fninf.2012.00007}, LONI~\cite{REX20031033}, 
        SPM~\cite{spm}, FastR~\cite{10.3389/fict.2016.00015},
        Automatic Analysis (AA)~\cite{10.3389/fninf.2014.00090} and 
        Pydpiper~\cite{10.3389/fninf.2014.00067}. These 
        workflow engines aim to satisfy four criteria: 1) simplified
        workflow composition, 2) performance, 3) portability of the workflows 
        and 4) reproducibility of the analyses. While these workflow engines
        do tackle performance, it is mainly limited to minimizing computation
        time and does not consider data transfer times -- which are typically
        costly in Big Data settings. Therefore, these engines need to be adapted
        for the processing of neuroimaging data.

        Big Data frameworks differ from those available in neuroimaging as their
        focus is mainly on workflow composition and performance. Big Data 
        frameworks such as Apache Spark~\cite{Zaharia:2016:ASU:3013530.2934664} 
        and Dask~\cite{rocklin2015dask} both aim to be easy-to-use,
        applicable to a wide variety of analyses and improve performance in the 
        processing of Big Data. These types of systems were primarily designed 
        to be efficient on commodity infrastructure in which transferring data
        over a network is extremely costly.

        Currently, there lacks a Big Data framework which is adapted for the
        processing of neuroimaging data. However, tools such as the Thunder 
        Project~\cite{Freeman:2014aa} allow for the processing of time-series
        neuroimaging data using Spark as the backend framework. Such tools
        provide built-in functions to process the data. For users to 
        extend functionality, it is necessary for them to understand the 
        backend framework. Other scientific workflow engines, such as 
        Toil~\cite{Vivian:2017aa} and Pegasus~\cite{DEELMAN201517}, have 
        incorporated features that enable more 
        performant workflow executions for scientific Big Data. They also enable
        the integration of Big Data frameworks as sub-pipelines within their 
        workflows. Despite integration with Big Data frameworks, it is not
        possible to create Big Data workflows using their API exclusively.

@@

        \tristan{Add a sentence to present the goal of the report.}
        This paper aims to evaluate the efficiency of current neuroimaging 
        frameworks in comparison to BigData frameworks and to determine the 
        extent to which BigData frameworks must be adapted to be effective
        neuroimaging workflow engines. 
        In section~\ref{datasets}, we will briefly introduce the two 
        types of datasets that are giving rise to the neuroimaging big 
        data. Section~\ref{platforms} will provide a brief introduction 
        to the data-sharing and processing platforms available to 
        researchers. We will conclude our 
        introduction on the differences between commodity and HPC infrastructure
        in section~\ref{infrastructure}.

        As we wish to evaluate and compare scientific workflow frameworks with
        Big Data frameworks for the processing of neuroimaging big data, we will
        break down our analyses by the four principal features necessary for
        effective neuroimaging workflow engines. The workflow engines discussed
        will include SPM, LONI, Nipype, PSOM, FastR, AA, Pydpiper 
        for neuroimaging; Taverna, Pegasus, Make, OpenMOLE, Nextflow, Galaxy, 
        Toil for other scientific workflows; and MapReduce, Spark and Dask for
        Big Data Frameworks. Chapter \ref{workcomp} will evaluate these 
        workflows on the basis of workflow composition: we will % 'in other words' is on my list of expressions to avoid
        examine the programming languages necessary to construct these workflows
        , how easy is it to add new tools and will these workflows be understood
        by the community. In Chapter \ref{performance}, we will proceed to 
        investigate the strategies employed by these frameworks to improve the 
        performance of their workflows. Chapter \ref{portability} will discuss
        the portability of the workflows written using these frameworks. 
        Workflow reproducibility will be then discussed in Chapter 
        \ref{reproducibility}.

@@
